Comparing Power Optimizations in the Reorder Buffer

### 2011-06-07 paper/paper.tex
Added more paper sections.

We model the input to our simulator as a stream of integer and floatpoint instructions.  The instructions are represented at the bit-level and model MIPS instructions as closely as possible.  Each instruction consists of 1) Instruction type, 2) Source and destination registers 3) a PC.  Table XX indicates the types of instructions modeled.  Each instruction type also has a predetermined cycle execution time designating the number of cycles that must elapse between instruction issue and the completion of the instruction.  As modeling functional units goes beyond the scope of this simulator, instructions are flagged as complete after the listed number of cycles has elapsed.  The actual cycle counts are the same as those modeled in MIT's Graphite many-core processor simulator[REFXX], with the exception of LOAD and STORE.  The cycle counts corresponding to LOAD and STORE instructions are estimated as a predetermined 20 cycles each.  This is simply an estimation as the actual cycle count is heavily dependent on cache architecture and benchmark specific cache behavior.  Since both of these topics are beyond the scope of the simulator, we estimate the completion time of LOAD and STORE instructions as a set value substantially longer than the other more accurate instruction types.
\begin{table*}
\centering
\begin{tabular}{|c|c|c|c|} \hline
Instruction Type&Cycle Count&Instruction Type&Cycle Count\\ \hline
\texttt{LOAD}&20&\texttt{STORE}&20\\ \hline
\texttt{ADD}&1&\texttt{FADD}&3\\ \hline
\texttt{SUB}&1&\texttt{FSUB}&3\\ \hline
\texttt{MULT}&3&\texttt{FMULT}&5\\ \hline
\texttt{DIV}&18&\texttt{FDIV}&6\\ \hline\end{tabular}
\caption{Modeled instruction types and their corresponding completion cycle counts.}
\end{table*}

@@

The remaining three instruction streams are modeled after SPEC-FP benchmarks.  The benchmarks were chosen to represent three major modes of operation:  floating point heavy execution, integer heavy execution, and mixed type execution.  For a floating point heavy execution, 173.applu was selected and modeled.  173.applu models parabolic/elliptical partial differential equations and features almost exclusively floating point operations as well as a large number of memory operations making it useful as both a floating point benchmark as well as a long running memory intensive benchmark.  191.fma3d is used to model integer heavy execution.  It models finite-element crash simulation and features very few floating point operations and substantially fewer memory operations than 173.applu making it an ideal light-weight integer workload.  Finally, 183.equake is used to represent the average workload.  It models seismic wave propagation and features both integer and floating point operations as well as moderate memory usage.  As such it acts as a more representative general workload than the original default workload.  Each of the SPEC-FP benchmarks feature a 6% register forwarding frequency as is reported by [XX paper ref].

For the purposes of our power comparisons, each of the above instruction workloads are normalized to generate 10000 instructions.  Such a large instruction window is ideal to allow buffers to fill and behavioral characteristics to develop.  The specific values used for each workload are listed in Table XX.
\begin{table*}
\centering
\begin{tabular}{|c|c|c|c|c|} \hline
ins{\char'137}profile&default{\char'137}profile&173{\char'056}applu&191{\char'056}fma3d&183{\char'056}equake\\ \hline
loadProc&0.10&0.32&0.12&0.37\\ \hline
storeProc&0.10&0.12&0.24&0.06\\ \hline
addProc&0.10&0.01&0.15&0.08\\ \hline
subProc&0.10&0.01&0.15&0.08\\ \hline
multProc&0.10&0.01&0.15&0.08\\ \hline
divProc&0.10&0.01&0.15&0.08\\ \hline
faddProc&0.10&0.13&0.01&0.07\\ \hline
fsubProc&0.10&0.13&0.01&0.06\\ \hline
fmultProc&0.10&0.13&0.01&0.06\\ \hline
fdivProc&0.10&0.13&0.01&0.06\\ \hline
forwardingRate&0.10&0.06&0.06&0.06\\ \hline
instCount&10000&10000&10000&10000\\ \hline\end{tabular}
\caption{Benchmark Instruction Profiles showing percent rate of occurrence of each instruction type.}
\end{table*}
\subsection{Model Implementation}
\subsubsection{Circular ROB}
\subsubsection{Distributed ROB}
\subsubsection{Dynamic ROB}
\subsubsection{Reduced Output Port ROB (Retention Latch)}
\subsection{Power Tabulation Implementation}
\section{Results}
hi
\subsection{subsection titl}
sub section.. with table!
\begin{table*}
\centering
\caption{Some Typical Commands}
\begin{tabular}{|c|c|l|} \hline
Command&A Number&Comments\\ \hline
\texttt{{\char'134}alignauthor} & 100& Author alignment\\ \hline
\texttt{{\char'134}numberofauthors}& 200& Author enumeration\\ \hline
\texttt{{\char'134}table}& 300 & For tables\\ \hline
\texttt{{\char'134}table*}& 400& For wider tables\\ \hline\end{tabular}
\end{table*}
% end the environment with {table*}, NOTE not {table}!

@@

%TODO add in bib

\end{document}

>>>>

\begin{table*}
\centering
\begin{tabular}{|c|c|c|c|} \hline
Instruction Type&Cycle Count&Instruction Type&Cycle Count\\ \hline
\texttt{LOAD}&20&\texttt{STORE}&20\\ \hline
\texttt{ADD}&1&\texttt{FADD}&3\\ \hline
\texttt{SUB}&1&\texttt{FSUB}&3\\ \hline
\texttt{MULT}&3&\texttt{FMULT}&5\\ \hline
\texttt{DIV}&18&\texttt{FDIV}&6\\ \hline\end{tabular}
\caption{Modeled instruction types and their corresponding completion cycle counts.}
\label{tab:instypetable}
\end{table*}

We model the input to our simulator as a stream of integer and floatpoint instructions.  The instructions are represented at the bit-level and model MIPS instructions as closely as possible.  Each instruction consists of 1) Instruction type, 2) Source and destination registers 3) a PC.  Table ~\ref{tab:instypetable} indicates the types of instructions modeled.  Each instruction type also has a predetermined cycle execution time designating the number of cycles that must elapse between instruction issue and the completion of the instruction.  As modeling functional units goes beyond the scope of this simulator, instructions are flagged as complete after the listed number of cycles has elapsed.  The actual cycle counts are the same as those modeled in MIT's Graphite many-core processor simulator\cite{graphite}, with the exception of LOAD and STORE.  The cycle counts corresponding to LOAD and STORE instructions are estimated as a predetermined 20 cycles each.  This is simply an estimation as the actual cycle count is heavily dependent on cache architecture and benchmark specific cache behavior.  Since both of these topics are beyond the scope of the simulator, we estimate the completion time of LOAD and STORE instructions as a set value substantially longer than the other more accurate instruction types.

@@

The remaining three instruction streams are modeled after SPEC-FP benchmarks \cite{rupnow}.  The benchmarks were chosen to represent three major modes of operation:  floating point heavy execution, integer heavy execution, and mixed type execution.  For a floating point heavy execution, 173.applu was selected and modeled.  173.applu models parabolic/elliptical partial differential equations and features almost exclusively floating point operations as well as a large number of memory operations making it useful as both a floating point benchmark as well as a long running memory intensive benchmark.  191.fma3d is used to model integer heavy execution.  It models finite-element crash simulation and features very few floating point operations and substantially fewer memory operations than 173.applu making it an ideal light-weight integer workload.  Finally, 183.equake is used to represent the average workload.  It models seismic wave propagation and features both integer and floating point operations as well as moderate memory usage.  As such it acts as a more representative general workload than the original default workload.  Each of the SPEC-FP benchmarks feature a 6{\char'045} register forwarding frequency as is reported by \cite{kucuk}.

\begin{table*}
\centering
\begin{tabular}{|c|c|c|c|c|} \hline
ins{\char'137}profile&default{\char'137}profile&173{\char'056}applu&191{\char'056}fma3d&183{\char'056}equake\\ \hline
loadProc&0.10&0.32&0.12&0.37\\ \hline
storeProc&0.10&0.12&0.24&0.06\\ \hline
addProc&0.10&0.01&0.15&0.08\\ \hline
subProc&0.10&0.01&0.15&0.08\\ \hline
multProc&0.10&0.01&0.15&0.08\\ \hline
divProc&0.10&0.01&0.15&0.08\\ \hline
faddProc&0.10&0.13&0.01&0.07\\ \hline
fsubProc&0.10&0.13&0.01&0.06\\ \hline
fmultProc&0.10&0.13&0.01&0.06\\ \hline
fdivProc&0.10&0.13&0.01&0.06\\ \hline
forwardingRate&0.10&0.06&0.06&0.06\\ \hline
instCount&10000&10000&10000&10000\\ \hline\end{tabular}
\caption{Benchmark Instruction Profiles showing percent rate of occurrence of each instruction type.}
\label{tab:insprofile}
\end{table*}

For the purposes of our power comparisons, each of the above instruction workloads are normalized to generate 10000 instructions.  Such a large instruction window is ideal to allow buffers to fill and behavioral characteristics to develop.  The specific values used for each workload are listed in Table ~\ref{tab:insprofile}.
\subsection{Model Implementation}
The following section details the model implementatations in our simulator.  The basic re-order buffer model we use as a base-line for comparison is a standard circular ROB model.  Section ~\ref{sec:circ} discusses its implementation.  We further model three optimizations for re-order buffers in an attempt to improve overall power usage: 1) Distributed (Cluster based) ROB, 2) Dynamically Sized ROB, 3) Reduced Output Port ROB (Retention Latch).  Each of these optimized models are built on top of the underlying circular ROB model to provide as close a behavioral and power comparison as possible.  Each of the optimizations, as well as the standard circular reorder buffer, are described in detail below.
\subsubsection{Circular ROB}
\label{sec:circ}
\subsubsection{Distributed (Cluster based) ROB}
\subsubsection{Dynamically Sized ROB}
\subsubsection{Reduced Output Port ROB (Retention Latch)}
\subsection{Power Tabulation Implementation}
\section{Results}
hi
\subsection{subsection titl}
sub section.. with table!
\begin{table*}
\centering
\caption{Some Typical Commands}
\begin{tabular}{|c|c|l|} \hline
Command&A Number&Comments\\ \hline
\texttt{{\char'134}alignauthor} & 100& Author alignment\\ \hline
\texttt{{\char'134}numberofauthors}& 200& Author enumeration\\ \hline
\texttt{{\char'134}table}& 300 & For tables\\ \hline
\texttt{{\char'134}table*}& 400& For wider tables\\ \hline\end{tabular}
\end{table*}
% end the environment with {table*}, NOTE not {table}!

@@

%TODO add in bib 
\begin{thebibliography}{------}
\bibitem{graphite}
	MIT Graphite Simulator
	https://github.com/mit-carbon/Graphite/wiki
\bibitem{kucuk}
Gurhan Kucuk, Dmitry Ponomarev, and Kanad Ghose. 2002. Low-complexity reorder buffer architecture. In Proceedings
of the 16th international conference on Supercomputing (ICS '02). ACM, New York, NY, USA, 57-66. DOI=10.
1145/514191.514202 http://doi.acm.org/10.1145/514191.514202
\bibitem{rupnow}
Kyle Rupnow, Arun Rodrigues, Keith Underwood, and Katherine Compton. 2006. Scientific applications vs. SPEC-FP: a comparison of program behavior. In Proceedings of the 20th annual international conference on Supercomputing (ICS '06). ACM, New York, NY, USA, 66-74. DOI=10.1145/1183401.1183413 http://doi.acm.org/10.1145/1183401.1183413 
\end{thebibliography}
\end{document}
